---
output:
  html_document: default
  pdf_document: default
---


# INSTRUCTIONS

-   This is a group assignment.
-   Submit your answer digitally as two files through Moodle:
    -   An R markdown file (extension **Rmd**). Use the template provided
        to you and provide your answers (both code and text) below each
        question.
    -   An **HTML** file &ldquo;knitted&rdquo; by RStudio including all the results and
        plots. More details on how to create these files were provided
        in class on week 3.
-   Follow the Style Guide (available on Moodle). You can be
    penalized on up to 20% in each question for which you do not follow
    the Style Guide.
-   Questions regarding the assignment should be posted <span class="underline">exclusively</span> on
    the respective discussion forum on Moodle.


## <span class="underline">Deadline:</span> Monday, March 10 at 23:59.

-   Late submissions are <span class="underline">not allowed</span>

<span class="underline">**Warning:**</span> The detection of <span class="underline">any form of plagiarism</span> in your work
means the assignment will be graded with <span class="underline">ZERO points</span>.

\newpage


# Dating Platforms

Online dating platforms enable people to find new connections with the
goal of developing personal, romantic, or sexual relationships. In
some platforms individuals express their interest in others in
multiple ways, including by &ldquo;liking&rdquo; them. &ldquo;Liking&rdquo; someone is an
indication of interest that, in most platforms, is revealed to both
individuals only if the &ldquo;liking&rdquo; is mutual, i.e., if the target
individual also stated interest. In such cases we say there is a
match.

For this assignment we are going to analyze &ldquo;likes&rdquo; issued by members
in an online dating platform using tools learned in class. Answer each
of the questions below the best you can. In case you find any
ambiguity in the question, assume the most sensible option in you
opinion, state your assumptions and proceed. In most cases ambiguity
is by design.


## Data Description

The file `nda-dating-likes.RData` contains two `data.table`, each of
them is described below.


### `dt.users`

<table border="2" cellspacing="0" cellpadding="6" rules="groups" frame="hsides">


<colgroup>
<col  class="org-left" />

<col  class="org-left" />

<col  class="org-left" />
</colgroup>
<thead>
<tr>
<th scope="col" class="org-left">Field</th>
<th scope="col" class="org-left">Type</th>
<th scope="col" class="org-left">Description</th>
</tr>
</thead>

<tbody>
<tr>
<td class="org-left"><code>user_id</code></td>
<td class="org-left">integer</td>
<td class="org-left">user identification</td>
</tr>


<tr>
<td class="org-left"><code>inviter_id</code></td>
<td class="org-left">integer</td>
<td class="org-left">id of the user that invited this user to the platform</td>
</tr>


<tr>
<td class="org-left"><code>gender</code></td>
<td class="org-left">string</td>
<td class="org-left">gender</td>
</tr>


<tr>
<td class="org-left"><code>birth_year</code></td>
<td class="org-left">integer</td>
<td class="org-left">year of birth</td>
</tr>


<tr>
<td class="org-left"><code>education</code></td>
<td class="org-left">integer</td>
<td class="org-left">education</td>
</tr>


<tr>
<td class="org-left"><code>approved_week</code></td>
<td class="org-left">date</td>
<td class="org-left">week in which this user was approved to the platform</td>
</tr>


<tr>
<td class="org-left"><code>height</code></td>
<td class="org-left">integer</td>
<td class="org-left">height (in cm)</td>
</tr>


<tr>
<td class="org-left"><code>children</code></td>
<td class="org-left">boolean</td>
<td class="org-left">whether the user has children</td>
</tr>


<tr>
<td class="org-left"><code>smoker</code></td>
<td class="org-left">boolean</td>
<td class="org-left">whether the user smokes</td>
</tr>


<tr>
<td class="org-left"><code>n_fb_friends</code></td>
<td class="org-left">integer</td>
<td class="org-left">how many Facebook friends does the user have</td>
</tr>
</tbody>
</table>


### `dt.likes`

<table border="2" cellspacing="0" cellpadding="6" rules="groups" frame="hsides">


<colgroup>
<col  class="org-left" />

<col  class="org-left" />

<col  class="org-left" />
</colgroup>
<thead>
<tr>
<th scope="col" class="org-left">Field</th>
<th scope="col" class="org-left">Type</th>
<th scope="col" class="org-left">Description</th>
</tr>
</thead>

<tbody>
<tr>
<td class="org-left"><code>sender_user_id</code></td>
<td class="org-left">integer</td>
<td class="org-left">user identification of the sender of the like</td>
</tr>


<tr>
<td class="org-left"><code>receiver_user_id</code></td>
<td class="org-left">integer</td>
<td class="org-left">user identification of the receiver of the like</td>
</tr>


<tr>
<td class="org-left"><code>week</code></td>
<td class="org-left">date</td>
<td class="org-left">week in which the like was issued</td>
</tr>
</tbody>
</table>

\newpage 


## Setup and data loading

Start by loading the required libraries and loading the likes data

```{r   }
library(data.table)
library(ggplot2)
library(igraph)

load("nda-dating-likes.RData")
```


## Questions


### Invites Network `[10 points]`

This section contains questions related to the invites network.

1.  *(Easy)* Build a directed graph representing the invites network: an
    individual A is connected to individual B if A invited B to the
    platform. What is the size of the longest chain? What is the
    clustering coefficient of this network? Justify. `[3 points]`

```{r}
# Vertices of the graph
all.user.ids <- unique(dt.users$user_id)

# Edges of the graph
dt.edges <- dt.users[!is.na(inviter_id), .(inviter_id, user_id)]

# Create the graph
g.invites.network <- graph_from_data_frame(d = dt.edges, directed = TRUE, vertices = all.user.ids)

# Plot the graph
plot(
  g.invites.network, 
  layout = layout_nicely(g.invites.network), 
  vertex.size = 3, 
  vertex.label = NA, 
  edge.arrow.size = 0.5
)

```


```{r}
# Size of the longest chain
length(get_diameter(g.invites.network, directed=TRUE))
```

```{r}
# Just for visualization purposes
# Calculate the diameter path vertices
diameter.path <- get_diameter(g.invites.network, directed = TRUE)

# Extract the subgraph based on the diameter path
g.subgraph.diameter <- induced_subgraph(g.invites.network, vids = V(g.invites.network)[diameter.path])

# Plot the subgraph with labels
plot(
  g.subgraph.diameter, 
  layout = layout_nicely(g.subgraph.diameter), 
  vertex.size = 20, 
  edge.arrow.size = 0.5, 
  vertex.label = V(g.subgraph.diameter)$name, 
  vertex.label.cex = 0.6
)
```

```{r}
# What is the clustering coefficient of this network?
transitivity(g.invites.network, type="global")
```

In order to build the network we have to be cautious with the NA values in the inviter_id column. We have to filter them out before creating the edges. We also have to create a list of all unique users for the vertices. This then results in a graph in which nodes with no inviter have no incoming edges but can have outgoing edges.
Due to the fact that a node that is already in the network cannot join again, the graph is acyclic, i.e. a DAG. Further, as a node can be invited at most once leading to a tree-like structure, in which the number of incoming arcs is <= 1 for all nodes. This means that the size of the longest chain is the length of the diameter, which is 4 edges long, i.e. traversing through 5 nodes. The clustering coefficient is 0, which means that the network is not very clustered. This is expected as, again, the network is acyclic and tree-like.


2.  *(Easy, once you get what you need to do)* Goel, Watts and Goldstein,
    in their paper entitled &ldquo;The structure of online diffusion
    networks&rdquo;, state that long cascades in diffusion networks happen
    very rarely in online networks. Do you think is also the case with
    the invites network in this online dating platform? Perform the
    analyses you deem necessary to answer this question with reasonable
    confidence. `[3 points]`
    
Three key observations: 
1. Small Cascade Sizes: The majority of diffusion cascades are small, often terminating within one degree of an initial adopting seed. 
2. Simple Tree Structures: A limited number of simple tree structures describe the vast majority of cascades, indicating that complex, branching cascades are rare.
3. Dominance of Few Individuals: Even in large cascades, a significant portion of adoptions occur close to a few dominant individuals, rather than being evenly distributed across many layers of referrals.

```{r}
# 1. Small Cascade Sizes:
# Get the connected components of the graph
connected.components <- components(g.invites.network)$membership

# Count the size of each connected component (cascade)
t.cascade.sizes <- table(connected.components)

# Plot the distribution of cascade sizes
ggplot(data.frame(size = as.numeric(t.cascade.sizes)), aes(x = size)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black", aes(y = after_stat(count))) +
  labs(title = "Distribution of Cascade Sizes",
       x = "Cascade Size",
       y = "Count")
```

```{r}
# Just for visualization purposes
# Find the largest cascade size
largest.cascade.size <- max(t.cascade.sizes)

# Find the connected component(s) with the largest cascade size
largest.cascade.components <- which(t.cascade.sizes == largest.cascade.size)

# Get the vertices belonging to the component with the largest cascade size
largest.cascade.vertices <- which(connected.components == largest.cascade.components[1])

# Extract the subgraph corresponding to the component with the largest cascade size
g.largest.cascade.subgraph <- induced_subgraph(g.invites.network, vid = largest.cascade.vertices)

# Plot the subgraph
plot(
  g.largest.cascade.subgraph, 
  layout = layout_nicely(g.largest.cascade.subgraph), 
  vertex.size = 3, 
  vertex.label = NA, 
  edge.arrow.size = 0.3
)
```

```{r}
# 2. Simple Tree Structures:
is_dag(g.invites.network)
```


```{r}
# 3. Dominance of Few Individuals:
# Histogram of the out-degree distribution
out.degrees <- degree(g.invites.network, mode = "out")
ggplot(data.frame(out.degree = out.degrees), aes(x = out.degree)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black", aes(y = after_stat(count))) +
  labs(title = "Out-Degree Distribution",
       x = "Out-Degree",
       y = "Count")
```
  
After looking at the three key observations from the paper, i.e. (1) Small Cascade Sizes, (2) Simple Tree Structures, and (3) Dominance of Few Individuals, we can conclude that the invites network in this online dating platform is consistent with the observations made by Goel, Watts and Goldstein. The distribution of cascade sizes shows that the majority of cascades are small, with the largest cascade being 336 nodes long. The network is acyclic and tree-like, which is consistent with the observation of simple tree structures. Finally, the out-degree distribution shows that the majority of nodes have an out-degree of 1, which is consistent with the observation of the dominance of few individuals.

3.  *(Moderate)* Do you think the invites network exhibits homophily?
    In other words, do you think men are more likely to invite other
    men and women more likely to invite other women? What about in
    terms of age? Do members tend to invite members of a similar age?
    Please describe your overall approach to answering these questions
    and justify and explain each step of your answer. `[4 points]`
    
```{r}

# Merge dt.users with itself to match inviter's gender with invitee's gender
dt.merged.gender <- merge(
  dt.users, 
  dt.users, 
  by.x = "inviter_id", 
  by.y = "user_id", 
  suffixes = c("_inviter", "_invitee")
)

# Count the number of invitees of the same gender for each inviter
dt.gender.distribution <- dt.merged.gender[, .N, by = .(inviter_id, gender_inviter, gender_invitee)]

# Reshape data to wide format
dt.gender.distribution <- data.table(
  dcast(
    dt.gender.distribution, 
    inviter_id + gender_inviter ~ gender_invitee, 
    value.var = "N", 
    fill = 0
  )
)

# Convert columns to numeric
dt.gender.distribution[, c("Male", "Female") := lapply(.SD, as.numeric), .SDcols = c("Male", "Female")]

# Calculate the probability of inviting an invitee of the same gender
dt.gender.distribution[, Probability := Male / (Male + Female)]

# Calculate average probabilities for male and female inviters
avg.prob.male <- mean(dt.gender.distribution[gender_inviter == "Male", Probability], na.rm = TRUE) * 100
avg.prob.female <- mean(dt.gender.distribution[gender_inviter == "Female", Probability], na.rm = TRUE) * 100

# Prepare data for heatmap
df.heatmap.data <- data.frame(
  inviter_gender = rep(c("Male", "Female"), each = 2),
  invitee_gender = rep(c("Male", "Female"), 2),
  probability = c(avg.prob.male, 100 - avg.prob.female,100 - avg.prob.male, avg.prob.female)
)

# Generate heatmap
ggplot(df.heatmap.data, aes(x = invitee_gender, y = inviter_gender, fill = probability)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "blue") +
  geom_text(aes(label = sprintf("%.2f%%", probability)), color = "black", size = 4) +
  labs(
    title = "Gender Homophily in Invitation Network", 
    x = "Invitee Gender", 
    y = "Inviter Gender", 
    fill = "Probability"
  ) +
  theme_minimal()

# Print the results
cat("Average probability of male inviter inviting a male invitee:", avg.prob.male, "\n")
cat("Average probability of female inviter inviting a female invitee:", avg.prob.female)
```

    
```{r}
library(reshape2)

# Calculate age based on birth year
dt.users[, Age := 2024 - birth_year]

# Bucketize age into ranges of 10
dt.users[, Age_bucket := cut(
  Age, 
  breaks = seq(from = 0, to = 140, by = 10), 
  labels = paste(seq(0, 130, by = 10), seq(9, 139, by = 10), sep = "-")
)]

# Merge dt.users with itself to match inviter's age with invitee's age
dt.merged.age <- merge(
  dt.users, 
  dt.users, 
  by.x = "inviter_id", 
  by.y = "user_id", 
  suffixes = c("_inviter", "_invitee")
)

# Ensure age bucket columns are character for accurate heatmap plotting
dt.merged.age[, Age_bucket_invitee := as.character(Age_bucket_invitee)]
dt.merged.age[, Age_bucket_inviter := as.character(Age_bucket_inviter)]

# Count invitations by age bucket combinations
dt.age.distribution <- dt.merged.age[, .(Count = .N), by = .(Age_bucket_inviter, Age_bucket_invitee)]

# Calculate total invitations by inviter age bucket
dt.age.distribution[, Total := sum(Count), by = Age_bucket_inviter]

# Calculate probability of inviting an invitee of a certain age bucket
dt.age.distribution[, Probability := Count / Total]

# Remove any rows with NA values to clean up the data
dt.age.distribution <- na.omit(dt.age.distribution)

# Reshape the data for heatmap visualization
dt.age.heatmap <- dcast(dt.age.distribution, Age_bucket_inviter ~ Age_bucket_invitee, value.var = "Probability")

# Melt the data for heatmap visualization
dt.age.heatmap <- melt(
  dt.age.heatmap, 
  id.vars = "Age_bucket_inviter", 
  variable.name = "Age_bucket_invitee", 
  value.name = "Probability"
)

# Plot the heatmap
ggplot(dt.age.heatmap, aes(x = Age_bucket_inviter, y = Age_bucket_invitee, fill = Probability)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  geom_text(
    aes(label = ifelse(is.na(Probability), "", sprintf("%.2f", Probability))), 
    color = "black", 
    size = 3
  ) +
  theme_minimal() +
  labs(
    x = "Inviter Age Bucket", 
    y = "Invitee Age Bucket", 
    fill = "Probability", 
    title = "Invitation Probability by Age Buckets"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```
```{r}
nrow(dt.users[gender == "Female"])/nrow(dt.users)
```

Within our predominantly female network, as female represent 55.2% of our users, we observe that males invite other males 70.47% of the time. This rate is notably more than what we would expect if invitations were randomly distributed based on gender. Female members exhibit a 19.35% probability of inviting other females, which indicates a relatively imbalanced invitation pattern. This suggests homophily among male user invitations, while it is the opposite for female.

Inspecting the heatmap for age-related invitation patterns reveals a strong but not exclusive preference for age homophily. High probabilities along the diagonal confirm that members tend to invite others of a similar age. However, the last column of the heatmap shows deviations from this pattern, indicating a lesser or varying degree of age homophily among the oldest age buckets.  It is important to account for such nuances, as they imply that while age homophily is a factor, other dynamics are also at play in shaping the invitation network.

4.  *(Make up question: You can skip any of the previous questions and
    answer this question instead.)* Describe and perform any additional
    analysis you want **using the invites network**. You will be graded
    by the creativity and correctness of the analysis, but also as a
    comparison to the best answers provided by other students. The best
    answers will have full points, while the other answers will be
    judged by comparison with the best answers. `[3 points]`
    
```{r}
# <answer here>
```
    
### Likes Network `[10 points]`

This section contains questions related to the likes network.

1.  *(Easy)* Build a directed graph representing the likes network: an
    individual A is connected to individual B if A liked B. What is the
    clustering coefficient of this network? Justify. `[2.5 points]`

```{r}
graph.likes <- graph_from_data_frame(dt.likes[, c("sender_user_id", "receiver_user_id")], directed = TRUE)

# For a dating platform, a low global clustering coefficient (0.0001293103) suggests users rarely form mutual liking triangles (A likes B, B likes C, C likes A), indicating sparse connections and few tight-knit groups. This highlights mainly one-way likes or separate user clusters, suggesting limited reciprocal attractions or distinct, non-interacting interest groups.


global.transitivity <- transitivity(graph.likes, type = "global")
print(global.transitivity)

```

```{r}

# On a dating platform, the local clustering coefficient indicates how users liking each others' profiles are interconnected, suggesting community-like structures. A higher coefficient means users interested in a profile also show interest in each other, forming tight-knit groups based on mutual attractions. This reflects not just individual connections but a web of mutual interests, hinting at shared preferences or compatibility within these user clusters.


local.clustering.coefficients <- transitivity(graph.likes, type = "local")
#print(local.clustering.coefficients) # hashed this line to spare extensive output

```


2.  *(Moderate)* How many individuals &ldquo;like&rdquo; individuals of the same
    gender? Does this happen more between women or between men? `[2.5 points]`



```{r}

# Add gender information for senders and receivers
likes.with.gender.sender <- merge(dt.likes, dt.users, by.x = "sender_user_id", by.y = "user_id", suffixes = c("", ".sender")) 
likes.with.gender.all <- merge(likes.with.gender.sender, dt.users, by.x = "receiver_user_id", by.y = "user_id", suffixes = c("", ".receiver"))

# Filter for same-gender likes
likes.same.gender <- likes.with.gender.all[likes.with.gender.all$gender == likes.with.gender.all$gender.receiver, ]

# INDIVIDUALS - Count unique individuals who like others of the same gender, by gender
unique.female.samegender <- length(unique(likes.same.gender$sender_user_id[likes.same.gender$gender == "Female"]))
unique.male.samegender <- length(unique(likes.same.gender$sender_user_id[likes.same.gender$gender == "Male"]))

# Print the number of unique individuals by gender
cat('Number of unique females who like others of the same gender:', unique.female.samegender, '\n')
cat('Number of unique males who like others of the same gender:', unique.male.samegender, '\n')

# Sum for total unique individuals
unique.total.samegender <- unique.female.samegender + unique.male.samegender
cat('Total number of unique individuals who like others of the same gender:', unique.total.samegender, '\n')

# TOTALS - Count number of same-gender like instances (not just unique users)
total.female.samegender.likes <- nrow(likes.same.gender[likes.same.gender$gender == "Female", ])
total.male.samegender.likes <- nrow(likes.same.gender[likes.same.gender$gender == "Male", ])

# Print the count of same-gender like instances
cat('Total instances of females liking other females:', total.female.samegender.likes, '\n')
cat('Total instances of males liking other males:', total.male.samegender.likes, '\n')

# Sum for total same-gender like instances
total.samegender.likes <- total.female.samegender.likes + total.male.samegender.likes
cat('Total number of same-gender like instances:', total.samegender.likes, '\n')

```

```{r} 

# Count of duplicate user_id
dt.duplicates <- dt.users[, .N, by = .(user_id)][N > 1]
count.duplicates <- nrow(dt.duplicates)
print(count.duplicates)

```

```{r}

# Count of different instances of genders
gender.count <- dt.users[, .(unique_users = uniqueN(user_id)), by = .(gender)]
print(gender.count)

```

```{r}

# Count sender gender 

count.sender.gender <- likes.with.gender.sender[, .N, by = .(gender)]
print(count.sender.gender)

```

The total number of likes sent between same genders, as well as the count of unique respective individuals shows hat both is more common among men, 58% and 57% respectively. Since there are 8,314 female and 6,746 male users, we cannot explain this based on the presence of the male gender. However, we can explain this with a higher share of male gender among total likes sent (67%). While there's an average of 255 likes per male user, the number for female users is much lower with an average of 105 likes.


3.  *(Easy)* Build an undirected graph representing the matches
    network: an individual A is connected to individual B if A liked B
    and B liked A. What is the clustering coefficient of this network?
    Justify. Hint: the function `as.undirected` transforms a directed
    graph into an undirected graph; it can transform only the mutually
    directed edges in undirected edges. `[2.5 points]`
    
```{r}
# Convert directed to undirected graph, keeping only mutual likes
g.matching <- as.undirected(graph.likes, mode = "mutual")

# Calculate global clustering coefficient of undirected graph
clustering.coefficient.mutual <- transitivity(g.matching, type = "global")

# Print the clustering coefficient
print(clustering.coefficient.mutual)

# The clustering coefficient of 3.226452e-05 the undirected graph, representing mutual likes in a dating network, indicates very sparse connections rather then the presence of thighly-knit communities. This low value suggests individuals' matches rarely match with each other, highlighting isolated interactions over communal or interconnected groupings. However, this is expectable as new users continuously join and new connections are formed dynamically. 
```

2nd approach to verify our results: Calculating clustering_coefficient Using "paste" check 

```{r}
# Find all unique pairs of senders and receivers (one)
unique.pairs <- unique(dt.likes[, c("sender_user_id", "receiver_user_id")])

# Identify mutual pairs with paste function - Check if each unique pair of users has a respective counter part
mutual.pairs <- unique.pairs[
  with(unique.pairs, paste(sender_user_id, receiver_user_id)) %in% 
    with(unique.pairs, paste(receiver_user_id, sender_user_id))
  , ]

# Remove one of the 2 mutual connections from mutual pairs
mutual.pairs.unique <- mutual.pairs[!duplicated(t(apply(mutual.pairs, 1, sort))), ]

# Creating an undirected graph from the unique dataframe
g <- graph_from_data_frame(mutual.pairs.unique, directed = FALSE)

# Calculating the clustering coefficient of the undirected graph
clustering.coefficient.mutual <- transitivity(g, type = "global")

# Printing clustering coefficient
print(clustering.coefficient.mutual)
```

4.  *(Moderate)* A recent research article gathered significant
    attention for claiming that, contrary to decades of prior research,
    most real life networks are not scale-free (i.e., the degree
    distribution does no not follow a power law). Based on the Quanta
    Magazine article (not on the original research paper) linked below,
    how do you classify the network of matches? Does it resemble a
    random network, a free-scale network, or a mixed network? Justify
    your answer. `[2.5 points]`
    
    Article: <http://bit.ly/2BwobR1>

In a scale-free network, we would expect some users to have a significantly higher number of matches than others, following a power-law distribution. This would mean that a few users would have many matches (hubs), while the majority of users would have only a few.

On the other hand, in a random network, each user would be equally likely to match with any other user, and the degree distribution (the number of matches per user) would tend to follow a Poisson distribution or a bell-shaped curve, where most users would have around the average number of matches, with few users having a very high or very low number of matches.

```{r}
#Proceeding to calculate degree frequency, global clustering coefficient and average path length

# Calculate and plot degree frequency of each node in matching graph
degree.distribution <- degree(g.matching)

# Calculate global clustering coefficient
clustering.coefficient.global <- transitivity(g.matching, type = "global")
print(paste("Global Clustering Coefficient:", clustering.coefficient.global))


# Calculate average path length
average.path.length <- mean_distance(g.matching)
print(paste("Average path length:", average.path.length))

# The dating platform's network, with its low clustering coefficient and deviation from a power-law degree distribution, doesn't align with typical scale-free or random network models. Its small average path length suggests some scale-free traits, indicating a unique, mixed network structure that reflects the complexity of real-world social connections.
```

```{r}
#Plotting Degree Frequency using mutual.pairs.unique
# Combining sender and receiver user IDs
all.users <- c(mutual.pairs.unique$sender_user_id, mutual.pairs.unique$receiver_user_id)

# Counting the number of matches (degree) for each user
degree.count <- table(all.users)

# Creating a frequency table for the degree distribution
degree.frequency <- table(degree.count)

# Converting to dataframe for plotting
degree.frequency.df <- as.data.frame(degree.frequency)

# Renaming columns appropriately
names(degree.frequency.df) <- c('Degree', 'Frequency')

# Plotting the degree distribution
library(ggplot2)
ggplot(degree.frequency.df, aes(x=Degree, y=Frequency)) +
  geom_bar(stat="identity") +
  theme_minimal() +
  labs(title="Degree Distribution of the Network",
       x="Degree", y="Frequency")
```

The network might be classified as scale-free because it seems like as if the data follows a power law, especially in the tail (a small number of users with a very high number of matches).

If the degree distribution would be close to a Poisson or normal distribution, it would suggest a random network.

However, in a scale-free network, the degree distribution follows a power law, meaning that the plot of frequency versus degree on a log-log scale would be a straight line, indicating that there are a few nodes with a very high degree (hubs) and many nodes with a low degree. The key feature of such networks is that the distribution has a heavy tail.

For a conclusive classification, we ideally fit a the graph on to a log-log scale.

```{r}
# Plotting on Log-Log Scale

# Ensuring there are no zero frequencies to avoid log(0)
degree.frequency.df.log <- subset(degree.frequency.df, Frequency > 0)

# Converting factors to numeric 
degree.frequency.df.log$Degree <- as.numeric(as.character(degree.frequency.df.log$Degree))
degree.frequency.df.log$Frequency <- as.numeric(as.character(degree.frequency.df.log$Frequency))

# Using ggplot2 to create a log-log plot
ggplot(degree.frequency.df.log, aes(x=Degree, y=Frequency)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  labs(title="Log-Log Degree Distribution of the Network",
       x="Degree (log scale)",
       y="Frequency (log scale)")
```

Heavy-tailed distribution: The distribution appears to have a heavy tail, where a small number of nodes have a very high degree compared to the rest. This is indicated by the data points that extend further to the right on the x-axis.

No straight line: A characteristic of a scale-free network is that the points on a log-log plot would approximate a straight line, since a scale-free network's degree distribution follows a power law. However, the data points in your plot show a curved relationship rather than a straight line, suggesting that the network may not be scale-free.

Possible cutoff: There appears to be a cutoff point (in the middle of the x-axis range) after which the frequency of nodes with higher degrees drops significantly. This could indicate a bounded scale-free network or a network that deviates from the scale-free model at higher degrees.

Plateauing for lower degrees: There seems to be a plateau or less steep decline in frequency for lower degree values, which deviates from what one would expect in a pure scale-free network.

Based on this plot alone, the network does not strictly follow a scale-free model since the points do not align in a straight line on a log-log scale. It more closely resembles either a mixed network or another type of network with its own unique degree distribution characteristics.


5.  *(Make-up question: You can skip any of the previous questions
    and answer this question instead.)* Describe and perform any
    additional analysis you want **using the any network originated from
    the likes data.** You will be graded by the creativity and
    correctness of the analysis, but also as a comparison to the best
    answers provided by other students. The best answers will have full
    points, while the other answers will be judged by comparison with
    the best answers. `[2.5 points]`
    

In order to run a successful business you need to understand your customer- / user-base. Thus we took a first simplified approach to categorize the users based on a User Engagement Score (UES): Total number of likes sent & received divided respectively by two.

```{r}
# Segmenting distinct user engagement levels

# Initialize dataframe with user_ids from graph
users.activity <- data.frame(user.ids = V(graph.likes)$name)  

# Extract out-degree (likes sent) and in-degree (likes received) to calculate UES
users.activity$out_degree <- degree(graph.likes, mode = "out")
users.activity$in_degree <- degree(graph.likes, mode = "in")

users.activity$UES <- (users.activity$out_degree + users.activity$in_degree) / 2

# Remove outliers based on nominal UES (top and bottom 5%)
UES.top.outlier <- quantile(users.activity$UES, 0.95)
UES.bottom.outlier <- quantile(users.activity$UES, 0.05)

users.filtered <- users.activity[users.activity$UES < UES.top.outlier & users.activity$UES > UES.bottom.outlier, ]

# Calculate UES range for filtered data
UES.min.filtered <- min(users.filtered$UES)
UES.max.filtered <- max(users.filtered$UES)
UES.range.filtered <- UES.max.filtered - UES.min.filtered  # Renamed for clarity

# Apply thresholds of 10% (by nominal UES) to categorize users 
active.filter <- UES.max.filtered - (0.1 * UES.range.filtered)
inactive.filter <- UES.min.filtered + (0.1 * UES.range.filtered)

# Correctly applying segment labels to the users.filtered data frame
users.filtered$segment <- ifelse(users.filtered$UES >= active.filter, 'Highly Active',
                                 ifelse(users.filtered$UES <= inactive.filter, 'Inactive', 'Passive'))

table(users.filtered$segment)
```

The segmentation reveals three distinct user engagement levels within the dating platform's likes network: 114 "Highly Active," 6,669 "Inactive," and 6,757 "Passive" users. The small number of highly active users are central to social interactions, indicating a core group driving engagement. The majority of users fall into the inactive and passive categories, highlighting a challenge in activating a significant portion of the user base. This distribution underscores the importance of strategies aimed at increasing engagement among less active users.
